\section{Phase 5: Supervised Learning \& Behavioral Cloning}
-

-
--
-
--
S T A R T  O F W O R K F L O W -
-
--
--

\noindent \textbf{Timeline: January 3 -- January 12, 2026}

The primary objective of Phase 5 was to synthesize a predictive model capable of replicating the \textbf{WiseX Agent's} decision policy with high fidelity. This phase documents the technical transition from a descriptive understanding of the Mexico City marketplace to the engineering of a "Predictive Brain" via a comparative algorithmic tournament.

\subsection{Feature Selection \& Data Regimes}
Prior to modeling, a dimensionality audit was conducted. While Principal Component Analysis (PCA) was utilized in the structural discovery phase, it served here to establish three distinct \textbf{Data Leagues} used to benchmark all contenders across varying levels of information density.

\subsubsection{The Tri-League Data Architecture}
Every algorithm in the tournament was evaluated across three specific data representations to isolate the impact of feature curation versus algorithmic power:

\begin{table}[H]
    \centering
    \small
    \begin{tabularx}{\textwidth}{l l X}
        \toprule
        \textbf{League} & \textbf{Description} & \textbf{Technical Purpose} \\
        \midrule
        \textbf{League A} & Unfiltered Set & Complete set of 65 raw features to test innate noise-filtration. \\
        \addlinespace
        \textbf{League B} & Curated Raw & The 20-feature "Praetorian Guard" set, utilizing raw scaled variables to satisfy the \textbf{Interpretability Mandate} (SHAP diagnostics). \\
        \addlinespace
        \textbf{League C} & Abstracted PCA & Orthogonalized 13-component set representing 90\% of system variance. \\
        \bottomrule
    \end{tabularx}
    \caption{Experimental Design: Data League definitions for the algorithmic tournament.}
    \label{tab:leagues}
\end{table}

\subsubsection{Categorical Feature Selection (CFS)}
High-value categorical features were engineered through three specific coalescence protocols to reduce cardinality while preserving strategic signal:
\begin{itemize}[noitemsep, topsep=3pt]
    \item \textbf{CFS1 (Geospatial Coalescence):} Merging hand-crafted heuristic polygons (Strategic Ground Truth) with HDBSCAN clusters (Machine Truth) to form a unified \texttt{zone\_id}.
    \item \textbf{CFS2 (Temporal Discretization):} Decomposition of timestamps into cyclic Hour/Day features and semantic "Day Blocks" (\textit{e.g., Morning Rush, Friday Gridlock}).
    \item \textbf{CFS3 (Contextual State):} Integration of product category, driver state, and the \texttt{heuristic\_flag} suite.
\end{itemize}

\subsection{The Algorithmic Tournament}
The tournament was executed across \textbf{17 discrete Jupyter Notebooks}, each representing a unique model architecture and environment trial.

\subsubsection{Contender Benchmarking}
Models were evaluated using \textbf{F1-Macro} scores to ensure performance was not inflated by the majority "Reject" class (which constitutes $\approx 92$\% of the observations). The tournament included:
\begin{itemize}[noitemsep]
    \item \textbf{Naive Bayes:} Evaluated in both Numerical (NB-Num) and Hybrid (NB-Cat) configurations.
    \item \textbf{Logistic Regression:} Benchmarked using \textit{Temporal Splits} to test chronological resilience and \textit{Stratified K-Fold} to ensure class representation.
    \item \textbf{Ensemble Methods:} Competitive evaluation of \textit{Random Forest} and \textit{XGBoost}.
\end{itemize}

\subsubsection{Hyperparameter Standardization}
To ensure valid comparative results, hyperparameters were held constant across "Leagues" within the same model family. This \textbf{Frozen Architecture} approach isolated the impact of the data representation, allowing for an empirical assessment of whether curated raw features or abstracted PCA components provided superior predictive signal.

\subsection{Champion Forensic: XGBoost \& The Gravitational Well}
\textbf{XGBoost} was identified as the tournament champion. However, a deep diagnostic audit of the initial monolithic confusion matrix revealed a critical failure mode: \textbf{The Gravitational Well}.

\medskip
\noindent \textbf{The Problem.} The model exhibited a "Lazy Attraction" toward the \texttt{dropoff\_non\_operational} majority class. Because geospatial rejections represent nearly 50\% of the dataset and follow deterministic boundaries, the model used them as a default attractor, effectively drowning out the strategic signals of nuanced classes such as the \textit{Expected Value Gamble} (EEUM).

\subsection{The Final Architecture: The Cognitive Cascade}
To neutralize the Gravitational Well effect, the project pivoted from a monolithic classifier to a \textbf{Hierarchical Cognitive Cascade}. This architecture decomposes the multi-class problem into a sequence of specialized inference layers:

\begin{itemize}[noitemsep]
    \item \textbf{Layer 1 (The Bouncer):} A binary filter isolating orthogonal, deterministic classes (Geospatial and Economic Exclusions).
    \item \textbf{Layer 2 (The Strategist):} A model operating exclusively on the high-yield subset. By removing majority-class noise, this layer is forced to learn the subtle distinctions between \textbf{Acceptance} and \textbf{Strategic Gambles} (EEUM).
\end{itemize}

\subsection{Technical Conclusion: The Temporal Wall}
The implementation of the Cognitive Cascade successfully isolated the strategic drivers but revealed a fundamental limit of classical Machine Learning in this domain: \textbf{Static Indifference}.

\subsubsection{The Snapshot Paradox (Statics vs. Dynamics)}
\textbf{The Constraint.} Classical ML architectures are inherently "Snapshot" engines. Even when enriched with stateful features (\textit{e.g., consecutive\_rejects}), the model treats each decision event as an isolated point in a vacuum.

\noindent \textbf{The Finding.} Analysis proved that \textit{Accepted} trips and \textit{EEUM} rejections occupy the same geometric space within a static tabular context. This occurs because the Agent’s decision to gamble is based on a **dynamic expectation of the stream's future yield**. While the Agent uses the "flow" to inform the bet, the model only sees the "row."

\subsubsection{The Deep Learning Pivot}
This finding provides the technical justification for the transition to \textbf{Phase 6}:
\begin{itemize}[noitemsep]
    \item \textbf{Sparsity Mitigation:} GANs are required to synthesize $1\text{M}+$ rows of market physics to overcome the $N=1$ data bottleneck.
    \item \textbf{Sequential Learning:} The synthetic manifold will be used to train \textbf{Recurrent Neural Networks (RNNs)} to learn the "flow" of the stream.
\end{itemize}

\subsubsection{Fleet-Scale Thesis}
For scalability, the project proposes an abstracted 3-category decision map: (1) \textbf{Accepted}, (2) \textbf{Rejected (EEUM)}, and (3) \textbf{Rejected (Other)}. This orthogonalization allows for high-frequency trading-style algorithms where the "Gamble" is treated as a formal financial derivative of search time.

\subsection*{Appendix: Binary Strategy Fallback}
Prior to the finalization of the multi-class cascade, a parallel \textbf{Binary Tournament (Accept vs. Reject)} was executed using NB, LogReg, and SVM. While achieving high global accuracy, these models were deprecated in favor of the Cognitive Cascade's ability to capture the nuance of strategic intent.



-

-
--
-
--
EE NN DD OOFF   O F W O R K F L O W -
-
--
--